```{r}
sys.source(file = 'R/descriptions/STH.R', envir = knitr::knit_global())
```

# Background

Herein, discuss helminthiases, and particularly focus on soil-transmitted helminth infections. School age children are children in the ae range 5 -- 14 years [@Montresor2011]

```{r}
terms <- STH()

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('species', 'name')),
    caption = 'The soil-transmitted helminth\\\\species in focus') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_footnote(label = '', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "p{0.18\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l")
```


Notes (Brooker 2006)

Environmental factors and prevalence spatial patterns
* temperature
* rainfall
* altitude
* soil moisture
* relative atmospheric humidity
* vegetation index

Under certain conditions eggs my not embryonate
* low humidity, $< 50%$: Ascaris lumbricoides, Trichuris trichiura

