
```{r introduction functions}
sys.source(file = '../R/metrics/DiseasePrevalence.R', envir = knitr::knit_global())
sys.source(file = '../R/metrics/DiseaseDALY.R', envir = knitr::knit_global())
```

```{r introduction settings}
pathstr <- dirname(getwd())
```


# *INTRODUCTION* \label{chapter:introduction}

Soil transmitted helminths (STH), or geohelminths, are parasitic intestinal worms that cause intestinal nematode infections (INI).  According to the Lancet's latest global burden of disease study (@Lancet2020, @IHME2020), as of 2019 INI - comprising *ascariasis*, *trichuriasis*, and *hookworm disease* - are the most prevalent of the neglected tropical diseases (\cref{tab:T0023}).  Additionally, INI are the second most burdensome neglected tropical diseases, after dengue (\cref{tab:T0024}).  Therefore, intestinal nematode infections are a focus of the World Health Organization (WHO).

\vspace{10pt}

```{r T0023}

terms <- DiseasePrevalence(pathstr = pathstr)

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    format.args = list(big.mark = ','),
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'prev.\n per\n $100k$', 'lower', 'upper')),
    caption = 'The five most prevalent neglected tropical diseases during the year 2019.  The unit of measure is prevalence per $100k$ inhabitants.  Data Source: \\href{https://vizhub.healthdata.org/gbd-compare/}{Institute for Health Metrics} $\\&$ \\href{https://www.thelancet.com/journals/lancet/issue/vol396no10258/PIIS0140-6736(20)X0042-0}{The Lancet Global Burden of Disease 2019}') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'prev.: prevalence', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "l")
```

\vspace{10pt}

Intestinal nematode infections prevention, control, and elimination activities depend on knowledge of prevalence (@Montresor2011), and prevalence metrics depend on field surveys, which are rather expensive to conduct.  They are especially expensive when there are no measures to systematically/strategically direct field scientists to survey areas (@Brooker2006).  In the absence of systematic/strategic direction, the field scientists have to survey as wide an area as possible, albeit subject to financial, accessibility, etc., constraints.

<br>

An option, in place of expensive field surveys, is INI prevalence predictions.  Prediction model development depends on probable risk factors knowledge & understandings (@Murray2020).   In relation to geohelminths infections, WaSH (water, sanitation, and hygiene) variables are amongst the sets of probable risk factors (@Albonico2006).  However, although there are a few studies concluding that one or more WaSH variables are geohelminths infections risk factors, the evidence is contentious (@Garn2022, @Freeman2017,  @Strunz2014).  This project investigates the possibility that one or more WaSH variables are geohelminths infections risk factors.

\vspace{10pt}

```{r T0024}
terms <- DiseaseDALY(pathstr = pathstr)

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    format.args = list(big.mark = ','),
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'DALY\n per\n $100k$', 'lower', 'upper')),
    caption = 'The five most burdensome neglected tropical diseases during the year 2019.  The unit of measure is disability adjusted life years (DALY) per $100k$ inhabitants, i.e., the number of years lost to disability, ill-health, or early death per $100k$ inhabitants.  Data Source: \\href{https://vizhub.healthdata.org/gbd-compare/}{Institute for Health Metrics} $\\&$ \\href{https://www.thelancet.com/journals/lancet/issue/vol396no10258/PIIS0140-6736(20)X0042-0}{The Lancet Global Burden of Disease 2019}') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'DALY: disability adjusted life years', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "l")

```

\vspace{10pt}

## *Project Aim*

Understanding the association between geohelminths infections prevalence and WaSH (water, sanitation, and hygiene) variables, and their effect, or otherwise, on geohelminths infections prevalence prediction.

## *Project Objectives*

Hence, the project's objectives are to

> 1. Understand the association between geohelminths infections prevalence and WaSH (water, sanitation, and hygiene) variables.
>
> 2. Determine the extent to which WaSH variables aid STH prevalence prediction

via geostatistical models.


## *Manuscript Structure*

The *background* chapter discusses the biology, ecology, and transmission dynamics of geohelminths; and a history of prevalence prediction. Next, the *data* chapter outlines the underlying data sets of this project.  This leads unto the *exploratory analysis* chapter, which explores the relationship, or otherwise, between the prevalence of a geohelminth infectious disease and independent variables. The *preliminary investigation* & *geostatistical binomial logistic models* chapters focus on explanatory modelling.  Finally, the *conclusion*/*discussion* chapter discusses the project's observations, findings, and limitations.

In this text intestinal nematode infections (INI) & geohelminth infections are synonyms, and the comprising diseases are *ascariasis*, *trichuriasis*, and *hookworm disease*.


