
<!--- functions --->
```{r introduction functions}
sys.source(file = '../R/metrics/DiseasePrevalence.R', envir = knitr::knit_global())
sys.source(file = '../R/metrics/DiseaseDALY.R', envir = knitr::knit_global())
sys.source(file = '../R/descriptions/DescriptionSpecies.R', envir = knitr::knit_global())
sys.source(file = '../R/metrics/MetricsSpecies.R', envir = knit_global())
```

<!--- settings --->
```{r introduction settings}
pathstr <- dirname(getwd())
```


# *INTRODUCTION* \label{chapter:introduction}

Soil transmitted helminths (STH), or geohelminths, are parasitic intestinal worms that cause intestinal nematode infections (INI).  According to the Lancet's latest global burden of disease study (@Lancet2020, @IHME2020), as of 2019 INI - comprising *ascariasis*, *trichuriasis*, and *hookworm disease* - are the most prevalent of the neglected tropical diseases (\cref{tab:T0023}).  Additionally, INI are the second most burdensome neglected tropical diseases, after dengue (\cref{tab:T0024}).  Therefore, intestinal nematode infections are a focus of the World Health Organization (WHO).

\vspace{10pt}

```{r T0023}

terms <- DiseasePrevalence(pathstr = pathstr)

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    format.args = list(big.mark = ','),
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'prev.\n per\n $100k$', 'lower', 'upper')),
    caption = 'The five most prevalent neglected tropical diseases during the year 2019.  The unit of measure is prevalence per $100k$ inhabitants.  Data Source: \\href{https://vizhub.healthdata.org/gbd-compare/}{Institute for Health Metrics} $\\&$ \\href{https://www.thelancet.com/journals/lancet/issue/vol396no10258/PIIS0140-6736(20)X0042-0}{The Lancet Global Burden of Disease 2019}') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'prev.: prevalence', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "l")
```

\vspace{10pt}

Intestinal nematode infections prevention, control, and elimination activities depend on knowledge of prevalence (@Montresor2011), and prevalence metrics depend on field surveys, which are rather expensive to conduct.  They are especially expensive when there are no measures to systematically/strategically direct field scientists to survey areas (@Brooker2006).  In the absence of systematic/strategic direction, the field scientists have to survey as wide an area as possible, albeit subject to financial, accessibility, etc., constraints.

<br>

An option, in place of expensive field surveys, is INI prevalence predictions.  Prediction model development depends on probable risk factors knowledge & understandings (@Murray2020).   In relation to geohelminths infections, WaSH (water, sanitation, and hygiene) variables are amongst the sets of probable risk factors (@Albonico2006).  However, although there are a few studies concluding that one or more WaSH variables are geohelminths infections risk factors, the evidence is contentious (@Garn2022, @Freeman2017,  @Strunz2014).  This project investigates the possibility that one or more WaSH variables are geohelminths infections risk factors.

\vspace{10pt}

```{r T0024}
terms <- DiseaseDALY(pathstr = pathstr)

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    format.args = list(big.mark = ','),
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'DALY\n per\n $100k$', 'lower', 'upper')),
    caption = 'The five most burdensome neglected tropical diseases during the year 2019.  The unit of measure is disability adjusted life years (DALY) per $100k$ inhabitants, i.e., the number of years lost to disability, ill-health, or early death per $100k$ inhabitants.  Data Source: \\href{https://vizhub.healthdata.org/gbd-compare/}{Institute for Health Metrics} $\\&$ \\href{https://www.thelancet.com/journals/lancet/issue/vol396no10258/PIIS0140-6736(20)X0042-0}{The Lancet Global Burden of Disease 2019}') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'DALY: disability adjusted life years', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "l")

```

\vspace{10pt}

## Geohelminths


An understanding of the biology, ecology, and transmission dynamics of geohelminths is key to proposing possible prevalence risk factors.  As previously noted, geohelminths infections comprises *ascariasis*, *trichuriasis*, and *hookworm disease*.  The underlying parasites are *A. lumbricoides*, *T. trichiura*, and *A. duodenale/N. americanus*, respectively (\cref{tab:T0001}).

\vspace{10pt}

```{r T0001}
terms <- DescriptionSpecies()

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'ascariasis', 'trichuriasis', '\\makecell[l]{hookworm\\\\disease}')),
    caption = 'The core geohelminths species that the World Health Organization focuses on.  In each case the infective stage organism lives in the environment/soil.') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_footnote(label = 'abbrv: abbreviation', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 4, latex_column_spec = "m{0.18\\\\linewidth}l", latex_valign = 'm')
```

\vspace{10pt}

During their adult stages the geohelminths parasites inhabit different parts of the human intestines, \cref{tab:T0001}, wherein they `reproduce sexually and produce eggs' [@Brooker2006]. The eggs exit humans via faeces, whilst defaecating, and mature in the environment.  Therein, they either (@Bethony2006, @Montresor2011)

> * Mature into infective eggs, which infect humans when they are accidentally ingested. This applies to A. lumbricoides and T. trichuria.

> * Hatch into infective larvae in the environment, soil.  The larvae subsequently infect humans by penetrating the skin.  This applies to N. americanus/A. duodenale.

These transmission dynamics illustrate why WaSH variables are plausible risk factors, and a few studies suggests evidence of WaSH interventions impacting **(a)** geohelminths infections prevalence and incidence, and **(b)** incidences of enteric infections, including geohelminth infections (@Serra2010, @Norman2010).  However, a few systematic reviews & meta-analyses suggests that the quality of the evidence is patchy (@Garn2022,  @Strunz2014, @Freeman2017).  The 2022 systematic review & meta-analysis by Garn et al. notes that

> *... The biological plausibility for improved access to WASH to interrupt transmission of STHs is clear, but WASH interventions as currently delivered have shown impacts that were lower than expected. There is a need for more rigorous and targeted implementation research and process evaluations in order that future WASH interventions can better provide benefit to users.*

`r tufte::quote_footer('--- Garn et al. [@Garn2022]')`

\vspace{10pt}

In the environment/soil a number of environmental factors aid or affect the survival of the pre-infective eggs, infective eggs (*A. lumbricoides*, *T. trichiura*), and infective larvae (*A. duodenale/N. americanus*).  Brooker et al. (@Brooker2006, @Brooker2000) review probable and known **(a)** environmental factors, and **(b)** ecological associations between geohelminths distributions and a variety of environmental factors.  The factors include temperature, soil moisture, relative atmospheric humidity, altitude, and more.  A rather critical observation in these reviews is that each species has a different, particular, spectrum of relationships with environmental factors.  Hence, this project focuses on a single disease only: Hookworm disease, as of $2019$ it is the most burdensome of the three diseases (\cref{tab:T0010}).

\vspace{10pt}

```{r T0010}
terms <- MetricsSpecies(pathstr = pathstr)

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'DALY\n per\n 100k', 'lower', 'upper', 'prev.\n per\n $100k$', 'lower', 'upper')),
    caption = 'The 2019 disability adjusted life years(DALY) per $100k$ inhabitants, and disease prevalence per $100k$ inhabitants.  Data Source: \\href{https://vizhub.healthdata.org/gbd-compare/}{Institute for Health Metrics} $\\&$ \\href{https://www.thelancet.com/journals/lancet/issue/vol396no10258/PIIS0140-6736(20)X0042-0}{The Lancet Global Burden of Disease 2019}') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2, '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'prev.: prevalence', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "l")
```

\vspace{10pt}

Altogether, the transmission dynamics suggests that WaSH factors are possible risk factors, whereas the ecological associations between geohelminths distributions and a variety of environmental factors suggests that a mix of climatic, topographic, and edaphic factors are also possible risk factors.  This project predominantly focuses on WaSH factors, this means - most probably - an insufficient collection of variables for a predictive model, but sufficient for **an explanatory model that models the association between a dependent variable and independent variables, whilst concurrently modelling missing knowledge**.


## Project Aim

Determine whether one or more WaSH (water, sanitation, and hygiene) variables are geohelminths infections prevalence risk factors.


## Project Objectives

Hence, the project's objectives are to

> 1. Determine which WaSH (water, sanitation, and hygiene) variables have a statistically significant impact on geohelminths infections prevalence.
>
> 2. Determine, estimate, the degree to which impacting WaSH variables affect geohelminths infections prevalence.

\noindent via geostatistical models.



## Manuscript Structure

The \hyperref[chapter:background]{background} chapter discusses the biology, ecology, and transmission dynamics of geohelminths. Next, the \hyperref[chapter:data]{data} chapter outlines the underlying data sets of this project.  This leads unto the \hyperref[chapter:exploration]{exploratory analysis} chapter, which explores the relationship, or otherwise, between the prevalence of a geohelminth infectious disease and independent variables. The \hyperref[chapter:preliminary]{preliminary investigation} & \hyperref[chapter:geo]{geostatistical binomial logistic models} chapters focus on explanatory modelling.  Finally, the [chapter:discussion]{discussion} chapter discusses the project's observations, findings, and limitations.

In this text intestinal nematode infections (INI) & geohelminths infections are synonyms, and the comprising diseases are *ascariasis*, *trichuriasis*, and *hookworm disease*.
