
# *INTRODUCTION* \label{chapter:introduction}

Soil transmitted helminths (STH), or geohelminths, are parasitic intestinal worms that cause intestinal nematode infections (INI).  According to the Lancet's latest global burden of disease study (@Lancet2020, @IHME2020), as of 2019 intestinal nematode infections - comprising *ascariasis*, *trichuriasis*, and *hookworm disease* - are the most prevalent of the neglected tropical diseases (\cref{fig:F0014}).  Additionally, INI are the second most burdensome neglected tropical diseases, after dengue (\cref{fig:F0015}).  Therefore, INI are a focus of the World Health Organization (WHO).

\vspace{10pt}

```{r F0014, fig.pos = 'h!', out.width = '85%', out.height = '85%', fig.align = 'center', fig.cap = 'The nine most prevalent neglected tropical diseases as of 2019.  Herein INI (Intestinal Nematode Infections) refers to the three core geohelminths infections: Ascariasis, Trichuriasis, and Hookworm Disease.'}
knitr::include_graphics(
  path = file.path('..', 'R', 'graphs', 'RatePrevalence.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\vspace{10pt}

Intestinal nematode infections prevention, control, and elimination activities depend on knowledge of prevalence [@Montresor2011], and prevalence metrics depend on field surveys, which are rather expensive to conduct.  They are especially expensive when there are no measures to systematically/strategically direct field scientists to survey areas [@Brooker2006].  In the absence of systematic/strategic direction, the field scientists have to survey as wide an area as possible, albeit subject to financial, accessibility, etc., constraints.

<br>

An option, in place of expensive field surveys, is INI prevalence predictions.  However, prediction models development depends on knowledge & understandings of probable risk factors [@Murray2020].   Accordingly, a key geohelminths infections prevalence hypothesis, of WHO experts, is that WaSH (water, sanitation, and hygiene) variables might be critical risk factors.  Hence, the project investigates the possibility that one or more WaSH variables are geohelminths infections risk factors that may aid geohelminths infections prevalence prediction.

\vspace{10pt}

```{r F0015, fig.pos = 'h!', out.width = '85%', out.height = '85%', fig.align = 'center', fig.cap = 'The nine most burdensome neglected tropical diseases as of 2019.  The DALY (diasbility adjusted life years) per 100k is the number of years lost to disability, ill-health, or early death per 100,000 inhabitants.  Again, INI (Intestinal Nematode Infections) refers to the three core geohelminths infections: Ascariasis, Trichuriasis, and Hookworm Disease.'}
knitr::include_graphics(
  path = file.path('..', 'R', 'graphs', 'RateDALY.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\vspace{10pt}

## *Project Aim*

Understanding the association between geohelminths infections prevalence and WaSH (water, sanitation, and hygiene) variables, and their effect, or otherwise, on geohelminths infections prevalence prediction.

## *Project Objectives*

Hence, the project's objectives are to

> 1. Understand the association between geohelminths infections prevalence and WaSH (water, sanitation, and hygiene) variables.
>
> 2. Determine the extent to which WaSH variables aid STH prevalence prediction

via geostatistical models.


## *Manuscript Structure*

The *background* chapter discusses the biology, ecology, and transmission dynamics of geohelminths; and a history of prevalence prediction. Next, the *data* chapter outlines the underlying data sets of this project.  This leads unto the *exploratory analysis* chapter, which explores the relationship, or otherwise, between the prevalence of a geohelminth infectious disease and independent variables. The *preliminary investigation* & *geostatistical binomial logistic models* chapters focus on explanatory modelling.  Finally, the *conclusion*/*discussion* chapter discusses the project's observations, findings, and limitations.

In this text intestinal nematode infections (INI) & geohelminth infections are synonyms, and the comprising diseases are *ascariasis*, *trichuriasis*, and *hookworm disease*.


