```{r L0003}
sys.source(file = '../R/descriptions/DescriptionSpecies.R', envir = knitr::knit_global())
```

# *BACKGROUND* \label{chapter:background}

\textcolor{duskyblue}{The background reading/research notes about}

* Strategies that have guided STH (soil transmitted helminth) field survey area selection.
* The biology, ecology, and transmission dynamics of geohelminths.

```{r T0001}
terms <- DescriptionSpecies()

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('species', 'name')),
    caption = 'The core geohelminths species that the World Health Organization focuses on.') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_footnote(label = '', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "p{0.18\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l")
```


\cref{tab:T0001} outlines the key geohelminths species [@Montresor2011].  There have been a variety of studies about the effects, or otherwise, of environmental factors, e.g., temperature, rainfall, altitude, soil moisture, relative atmospheric humidity, etc., on geohelminths infections prevalence patterns [@Brooker2006] ...
