```{r L0003}
sys.source(file = '../R/descriptions/DescriptionSpecies.R', envir = knitr::knit_global())
```

# *BACKGROUND* \label{chapter:background}

An understanding of the biology, ecology, and transmission dynamics of geohelminths is key to positing possible prevalence risk factors.  As noted earlier, the geohelminths in focus are *Ascaris lumbricoides*, *Trichuris trichuria*, *Necator americanus / Ancylostoma duodenale* (*Table ...*).  During their adult stages these geohelminths inhabit different parts of the human intestines (*Table ...*) wherein they ``reproduce sexually and produce eggs'' [@Brooker2006]. These eggs exit humans via faeces, whist defaecating.  The eggs mature in the environment [@Montresor2011], and they either

> * mature into infective eggs, which infect humans when they are accidentally ingested: Ascaris lumbricoides (roundworm) & Trichuris trichuria (whipworm) [@Brooker2006]
>
> * hatch into infective larvae in the environment, soil, which subsequently infect humans by penetrating the skin: Necator americanus / Ancylostoma duodenale (hookworm) [@Bethony2006]

Considering the transmission dynamics, the WaSH variables hypothesis is understandable.  A number  of factors aid the survival of the eggs, infective eggs, and infective larvae in the environment ...


```{r T0001}
terms <- DescriptionSpecies()

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('species', 'name')),
    caption = 'The core geohelminths species that the World Health Organization focuses on.') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_footnote(label = '', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "p{0.18\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l")
```


In line with the World Health Organization's WaSH hypothesis, the studies by Mascarini-Serra et al. [@Serra2010],

Altogether, the transmission dynamics suggests that WaSH factors are possible risk factors, whereas the ecology of free-living eggs - infective & pre-infective - and free-living infective larvae hint at possible environmental & climatic risk factors.^[Free-living: living in the environment, outwith the human host.]

This project predominantly focuses on WaSH factors, this means - most probably - an insufficient collection variables for a predictive model, but sufficient for an explanatory model that models/captures the missing knowledge.

Because WaSH, climatic, geographic, etc., factors have rather different effects on each parasite's life cycle stages, this project focuses on a single parasite class: Hookworm.  It is the most burdensome of the three [Table: Prevalence & DALY]