
# *EXPLORATORY ANALYSIS* \label{chapter:exploration}

This project's underlying data set is the Togo data set; it is the best vis-&agrave;-vis minimal missing points, missing data pattern, and the number of observations.

```{r T0007}
T <- read.csv(file = file.path(dirname(getwd()), 'warehouse', 'reference', 'frequencies', 'timeFrequencies.csv'))

T %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('year', 'N')),
    caption = 'The number of geohelminths infections examinations records per year. N denotes number of records.') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "m{0.235\\\\linewidth}l", latex_valign = 'm')

```

## The Distribution of Examinations Sites

```{r F0001, fig.pos = 'h!', out.extra = 'trim=5 135 15 135,clip', out.height = '42.5%', dpi = 175, fig.align = 'center', fig.cap = 'The distribution of geohelminths infections examinations sites in Togo; only 4 sites were examinations sites during both years.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'reference', 'distributions', 'mapDistributions.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\cref{fig:F0001} illustrates the distribution of examination sites across Togo.  The years $2009$ & $2015$ only have $4$ examination sites in common.  This observation has a bearing on the Preliminary Investigation chapter (\Cref{chapter:preliminary}), which informs geostatistical modelling. \textcolor{duskyblue}{[Discuss the rationale for using the data of the year 2015.]}


\vspace{25pt}

## Prevalence Distributions

```{r F0002, fig.pos = 'h!', out.width = '65%', fig.align = 'center', fig.cap = 'The distribution of geohelminths infections prevalence during the years 2009 and 2015'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'reference', 'distributions', 'densityDistributions.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\vspace{25pt}

## WaSH

```{r F0003, fig.pos = 'h!', out.height = '85%', fig.align = 'center', fig.cap = 'The relationships between Hookworm disease infections prevalence and access to sewage facilities$\\;$ an access fraction of 1 means $100\\%$ of the population has access.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'reference', 'sewer', 'disaggregateSewerReal.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\clearpage

## Elevation & Population Density

```{r F0007, fig.pos = 'h!', out.height = '85%', fig.align = 'center', fig.cap = 'The relationships between Hookworm disease infections prevalence and (a) elevation, (b) population density.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'reference', 'miscellaneous', 'disaggregateMiscellaneousReal.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```
























