

# *GEOSTATISTICAL BINOMIAL LOGISTIC MODELS* \label{chapter:geo}

The predominant focus of this chapter is explanatory models.  An explanatory model determines which independent variables have a statistically significant impact on a dependent variable in question, it also outlines the degree of impact of the independent variables.  Hence, and in relation to the project's aim & objectives, the objective herein is to

\begin{minipage}[c]{0.9\textwidth}

\vspace{15pt}

\textit{Determine which extraneous independent variables -  WaSH variables, population density, and elevation - have a statistically significant impact on the prevalence of Hookworm disease infections, and outline their degree of impact.}

\vspace{15pt}

\end{minipage}

Leading on from the \hyperref[chapter:data]{data}, \hyperref[chapter:exploration]{exploratory analysis}, \hyperref[chapter:preliminary]{preliminary investigation} chapters, a few important points:  **(a)** The outcomes - prevalence - are proportions, hence the project's focus is Binomial models. **(b)** The preliminary investigation identified Binomial generalised linear mixed models with strong evidence of residual spatial correlation, hence \textit{this chapter focuses on geostatistical binomial logistic [explanatory] models}. **(c)** Modelling is via the data set of a single year; Togo, $2015$.


## Geo-statistical binomial logistic models

Here we explore *geostatistical binomial logistics models* wherein the parameter estimates are either (a) Monte Carlo Maximum Likelihood estimates, or (b) Bayesian estimates.  The model structure is

\begin{align}
\renewcommand{\arraystretch}{1.5}
Y_{i} & \sim Binomial(e(x_{i}), p(x_{i})) (\#eq:0037) \\
\eta(x_{i}) & = ln \left( \frac{p(x_{i})}{1 - p(x_{i})} \right) (\#eq:0038) \\
& = \mathbf{\delta}(x_{i})^{T}\mathbf{\beta} + S(x_{i}) + U_{i} (\#eq:0039)
\end{align}

\vspace{10pt}

Again, \cref{tab:T0004} defines most terms.  $x_{i}$ is the spatial/geographic location of the $i^{th}$ observation.   $\mathbf{\delta}(x_{i})$ is the $i^{th}$ vector of fixed terms - i.e., the $i^{th}$ row vector of the model's design matrix - it depends on zero or more independent variables. Whilst $\mathbf{\beta}$ is the vector of the concurrent fixed effects coefficients.  $S(x_{i})$ is the stochastic process term, it models the *unexplained spatial variation* (@Diggle2019).  Herein, the assumption is that $S(\mathbf{x})$ is

> * A zero-mean gaussian process, i.e., the joint probability of ${S(x_{1}), \ldots, S(x_{n})}$ is multi-variate normal.\newline
> * Stationary, i.e., it has a constant variance - denoted $\sigma^{2}$.\newline
> * Isotropic, i.e., the correlation between $S(x_{i})$ & $S(x_{j})$ depends on the distance between $x_{i}$ & $x_{j}$ only.\newline

A consequence of these assumptions is that the covariance $\gamma(x_{i}, x_{j})$ is defined as
\begin{equation}
\gamma(x_{i}, x_{j}) = \sigma^{2}\rho(\left\Vert x_{i} - x_{j} \right\Vert) (\#eq:0025)
\end{equation}
wherein $\rho(u) = \rho(\left\Vert x_{i} - x_{j} \right\Vert)$ is the correlation function; $u = \left\Vert \cdot \right\Vert$ denotes distance.  Throughout this paper correlation function in use is the *exponential correlation function* (@Diggle2019)
\begin{equation}
\rho(u; \phi) = e^{-u/\phi}, \qquad u \geq 0 (\#eq:0016)
\end{equation}
The term $\phi$ is the scale parameter.

\vspace{10pt}

In contrast to $S(x_{i})$, $U_{i}$ are independent Normally distributed random variables
\begin{equation}
U_{i} \sim \mathcal{N}(0, \tau^{2}) (\#eq:0023)
\end{equation}
that capture the effects of unknown/unmeasured independent variables, i.e., $U_{i}$ represent the *unstructured variation
in $Y_{i}$* (@Diggle2019).  The unknown variables either **(a)** lack a spatial structure, or **(b)** *``... are spatially
varying but on a scale smaller than the minimum observed distance.''* (@Diggle2019)


### Linear Predictors

The \hyperref[chapter:preliminary]{preliminary investigation} chapter identifies two generalised linear mixed models whose **(a)** residuals exhibit spatial correlation, and **(b)**  coefficient estimates $\beta_{j}$ are significant $\forall \; j$, at significance level $0.05$.  The linear predictors of these models are

\begin{align}
\eta_{i} & = \color{darkolive}{\beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{piped\underline{\hspace{0.125cm}}sewer}(x_{i})^{2}} \nonumber \\
& \quad  \color{darkolive}{\; + \; \beta_{3}{elevation.km}(x_{i})} \color{black}{\; + \; b_{i}} \nonumber
\end{align}
and
\begin{align}
\eta_{i} & = \color{darkolive}{\beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{ln(p\underline{\hspace{0.125cm}}density.k)}(x_{i})} \nonumber \\
& \quad  \color{darkolive}{\; + \; \beta_{3}{elevation.km}(x_{i})} \color{black}{\; + \; b_{i}} \nonumber
\end{align}

Each linear predictor's fixed-term expression is the olive coloured part of the linear predictor.  Hence, and building on these preliminary investigation fixed-term expressions, the geostatistical binomial logistic models in focus are the models with linear predictors
\begin{align}
\eta_{i} & = \beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{piped\underline{\hspace{0.125cm}}sewer}(x_{i})^{2} (\#eq:0040) \\
& \quad  + \beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}\nonumber
\end{align}
and
\begin{align}
\eta_{i} & = \beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{ln(p\underline{\hspace{0.125cm}}density.k)}(x_{i}) (\#eq:0041) \\
& \quad  + \beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i} \nonumber
\end{align}

The next few sections discuss the models' estimates, and whether the models successfully capture **(a)** unexplained spatial variation, and **(b)** unstructured variation in $Y_{i}$.



\clearpage

## The Estimated Coefficients

\cref{tab:T0011} & \cref{tab:T0012} outline the coefficient estimates vis-&agrave;-vis the fixed-terms of linear predictors \cref{eq:0040} & \cref{eq:0041}, respectively.    By the hypotheses
\begin{minipage}[c]{0.9\textwidth}
$$H_{0}: \beta_{j} = 0$$
$$H_{1}: \beta_{j} \neq 0$$
\end{minipage}
\vspace{10pt}
\noindent each fixed-term coefficient estimate of \cref{tab:T0011} is statistically significant at the $0.05$ significance level, i.e., there is evidence in favour of the alternative hypothesis, $H_{1}$, $\forall \; j$, at significance level $0.05$.  These \cref{tab:T0011} observations are in-line with the \hyperref[section:preliminary-coefficients]{preliminary investigation's estimated coefficients} observations.

In contrast, not all the fixed-term coefficient estimates of \cref{tab:T0012} are significant; the coefficients of the $piped\underline{\hspace{0.125cm}}sewer$ & $ln(p\underline{\hspace{0.125cm}}density.k)$ terms are not statistically significant at the $0.05$ significance level.

\textbf{Hence, the rest of this chapter discusses the estimates of the Binomial geostatistical logistic model with linear predictor} \cref{eq:0040}.

\vspace{10pt}

<!--- coefficients --->
```{r T0011, fig.pos = 'h!'}
parabolic <- readRDS(file = file.path(dirname(getwd()), 'warehouse', 'models', 'nugget', 'blm', 'A', 'estimations.rds'))


parabolic$coefficients %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'coef.', 'coef.\nest.',
                                        'lower', 'upper', 's.e.', '$p_{value}$')),
    caption = 'The estimated coefficients of the Binomial geostatistical logistic model with linear predictor $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i})^{2} + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', '', 'confidence\ninterval' = 2, '', '')) %>%
  kableExtra::add_footnote(label = c('coef.: coefficient', 'est.: estimate', 's.e.: standard error'), notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```

<!--- coefficients --->
```{r T0012, fig.pos = 'h!'}
logarithmic <- readRDS(file = file.path(dirname(getwd()), 'warehouse', 'models', 'nugget', 'blm', 'B', 'estimations.rds'))


logarithmic$coefficients %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'coef.', 'coef.\nest.',
                                        'lower', 'upper', 's.e.', '$p_{value}$')),
    caption = 'The estimated coefficients of the Binomial geostatistical logistic model with linear predictor $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{ln(p\\underline{\\hspace{0.125cm}}density.k)}(x_{i})  + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', '', 'confidence\ninterval' = 2, '', '')) %>%
  kableExtra::add_footnote(label = c('coef.: coefficient', 'est.: estimate', 's.e.: standard error'), notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```

\clearpage

## The Variance & Scale Estimates

\cref{tab:T0025} outlines the variance, and scale parameter, estimates of linear predictor \cref{eq:0040}.  The confidence interval of the scale parameter $\phi$ is rather wide.  On the other hand, $\sigma^{2}$, the variance of the stochastic process $S(\mathbf{x})$, has a rather narrow confidence interval.

<!--- variance & scale estimates  --->
```{r T0025}
parabolic$special %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'est.', 'lower', 'upper', '$e^{est.}$', 'lower', 'upper')),
    caption = 'The Binomial geostatistical logistic model variance estimates vis-\\`{a}-vis $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i})^{2} + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$.  $\\kappa = 0.5$') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 8, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', 'confidence\ninterval' = 2, '', 'confidence\ninterval' = 2)) %>%
  kableExtra::add_footnote(label = 'est.: estimate', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, latex_column_spec = "c")
```


## Residual Spatial Correlation

The purpose of the Binomial geostatistical logistic model's [linear predictor] stochastic term $S(x_{i})$ is the modelling/capturing of *unexplained spatial variation* (@Diggle2019).  If the $S(x_{i})$ term of \cref{eq:0040} successfully models the unexplained spatial variation, then the associated standardised residual errors will be free of residual spatial correlation.  As before, we test for residual spatial correlation via the empirical variogram.  Foremost, the calculation of the standardised residual errors.

\vspace{10pt}

The standardised residual errors are
\begin{align}
r_{i} & = \frac{1}{\psi} \epsilon_{i}\left(1 - h_{i}\right)^{-1/2}, \quad \text{\textit{wherein}}  (\#eq:0010) \\
\psi^{2} & = \frac{1}{n - \rho} \sum^{n}_{i = 1}{\epsilon^{2}_{i}}   (\#eq:0011)
\end{align}
The number of observations is $n$, $\rho$ is the number of fixed effects parameters, i.e., the number of model coefficients.  The raw residuals are ${\epsilon_{1}, \ldots, \epsilon_{n}}$; each $\epsilon_{i}$ is the difference between observed outcome $Y_{i}$ and its fitted value. $\psi$ is the standard error of the raw residuals.  Finally, $h_{i}$ is the leverage of the $i^{th}$ observation.  The definition of leverage value is
\begin{equation}
\mathbf{h} = diag \Bigl( X\bigl(X^{T}X\bigr)^{-1}X^{T}  \Bigr) (\#eq:0012)
\end{equation}
wherein $X$ denotes a model's design matrix.

\vspace{10pt}

\cref{fig:F0008} outlines the empirical variogram of the standardised residual errors of the Binomial geostatistical logistic model with linear predictor \cref{eq:0040}.  The empirical variogram does not exhibit residual spatial correlation, i.e., the Binomial geostatistical logistic model in question successfully models the unexplained spatial variation.

<!--- empirical variogram --->
```{r F0008, fig.pos = 'h!', out.width = '60%', fig.align = 'center', fig.cap = 'The variogram of the Binomial geostatistical logistic model with linear predictor $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i})^{2} + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'A', 'variogram.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

\clearpage

## Prevalence Estimates

The project set aside a portion of Togo's $2015$ data set for preliminary prediction analysis purposes.  Thus far, the portion of the Togo $2015$ data set in use is, more or less, the *training* data set.  

<!--- prevalence --->
```{r F0009, fig.pos = 'h!', out.width = '85%', fig.align = 'center', fig.cap = 'The prevalence estimates and originals vis-\\`{a}-vis the Binomial geostatistical logistic model with linear predictor $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i})^{2} + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'A', 'diagonal.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```




### Bias & RMSE

<!--- bias & rmse --->
```{r T0013}

parabolic$discrepancies %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'Bias', 'RMSE')),
    caption = 'The bias and root mean square error values vis-\\`{a}-vis the Binomial geostatistical logistic model with linear predictor $\\beta_{0} + \\beta_{1}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i}) + \\beta_{2}{piped\\underline{\\hspace{0.125cm}}sewer}(x_{i})^{2} + \\beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$.') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```
