

# *GEOSTATISTICAL BINOMIAL LOGISTIC MODELS* \label{chapter:geo}

The predominant focus of this chapter is explanatory models.  An explanatory model determines which independent variables have a statistically significant impact on a dependent variable in question, it also outlines the degree of impact of the independent variables.  Hence, and in relation to the project's aim & objectives, the objective herein is to

\begin{minipage}[c]{0.9\textwidth}

\vspace{15pt}

\textit{Determine which extraneous independent variables -  WaSH variables, population density, and elevation - have a statistically significant impact on the prevalence of Hookworm disease infections, and outline their degree of impact.}

\vspace{15pt}

\end{minipage}

Leading on from the \hyperref[chapter:data]{data}, \hyperref[chapter:exploration]{exploratory analysis}, \hyperref[chapter:preliminary]{preliminary investigation} chapters, a few important points:  **(a)** The outcomes - prevalence - are proportions, hence the project's focus is Binomial models. **(b)** The preliminary investigation identified Binomial generalised linear mixed models with strong evidence of residual spatial correlation, hence \textit{this chapter focuses on geostatistical binomial logistic [explanatory] models}. **(c)** Modelling is via the data set of a single year; Togo, $2015$.


## Geo-statistical binomial logistic models

Here we explore *geostatistical binomial logistics models* wherein the parameter estimates are either (a) Monte Carlo Maximum Likelihood estimates, or (b) Bayesian estimates.  The model structure is

\begin{align}
\renewcommand{\arraystretch}{1.5}
Y_{i} & \sim Binomial(e(x_{i}), p(x_{i})) (\#eq:0037) \\
\eta(x_{i}) & = ln \left( \frac{p(x_{i})}{1 - p(x_{i})} \right) (\#eq:0038) \\
& = \mathbf{\delta}(x_{i})^{T}\mathbf{\beta} + S(x_{i}) + U_{i} (\#eq:0039)
\end{align}

\vspace{10pt}

Again, \cref{tab:T0004} defines most terms.  $x_{i}$ is the spatial/geographic location of the $i^{th}$ observation.   $\mathbf{\delta}(x_{i})$ is the $i^{th}$ vector of fixed terms - i.e., the $i^{th}$ row vector of the model's design matrix - it depends on zero or more independent variables. Whilst $\mathbf{\beta}$ is the vector of the concurrent fixed effects coefficients.  $S(x_{i})$ is the stochastic process term, it models the *unexplained spatial variation* (@Diggle2019).  Herein, the assumption is that $S(\mathbf{x})$ is

> * A zero-mean gaussian process, i.e., the joint probability of ${S(x_{1}), \ldots, S(x_{n})}$ is multi-variate normal.\newline
> * Stationary, i.e., it has a constant variance - denoted $\sigma^{2}$.\newline
> * Isotropic, i.e., the correlation between $S(x_{i})$ & $S(x_{j})$ depends on the distance between $x_{i}$ & $x_{j}$ only.\newline

A consequence of these assumptions is that the covariance $\gamma(x_{i}, x_{j})$ is defined as
\begin{equation}
\gamma(x_{i}, x_{j}) = \sigma^{2}\rho(\left\Vert x_{i} - x_{j} \right\Vert) (\#eq:0025)
\end{equation}
wherein $\rho(u) = \rho(\left\Vert x_{i} - x_{j} \right\Vert)$ is the correlation function; $u = \left\Vert \cdot \right\Vert$ denotes distance.  Throughout this paper correlation function in use is the *exponential correlation function* (@Diggle2019)
\begin{equation}
\rho(u; \phi) = e^{-u/\phi}, \qquad u \geq 0 (\#eq:0016)
\end{equation}
The term $\phi$ is the scale parameter.

\vspace{10pt}

In contrast to $S(x_{i})$, $U_{i}$ are independent Normally distributed random variables
\begin{equation}
U_{i} \sim \mathcal{N}(0, \nu^{2}) (\#eq:0023)
\end{equation}
that capture the effects of unknown/unmeasured independent variables, i.e., $U_{i}$ represent the *unstructured variation
in $Y_{i}$* (@Diggle2019).  The unknown variables either **(a)** lack a spatial structure, or **(b)** *``... are spatially
varying but on a scale smaller than the minimum observed distance.''* (@Diggle2019)


### Linear Predictors

The \hyperref[chapter:preliminary]{preliminary investigation} chapter identifies two generalised linear mixed models whose **(a)** residuals exhibit spatial correlation, and **(b)**  coefficient estimates $\beta_{j}$ are significant $\forall \; j$, at significance level $0.05$.  The linear predictors of these models are

\begin{align}
\eta_{i} & = \color{darkolive}{\beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{piped\underline{\hspace{0.125cm}}sewer}(x_{i})^{2}} \nonumber \\
& \quad  \color{darkolive}{\; + \; \beta_{3}{elevation.km}(x_{i})} \color{black}{\; + \; b_{i}} \nonumber
\end{align}
and
\begin{align}
\eta_{i} & = \color{darkolive}{\beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{ln(p\underline{\hspace{0.125cm}}density.k)}(x_{i})} \nonumber \\
& \quad  \color{darkolive}{\; + \; \beta_{3}{elevation.km}(x_{i})} \color{black}{\; + \; b_{i}} \nonumber
\end{align}

Each linear predictor's fixed-term expression is the olive coloured part of the linear predictor.  Hence, and building on these preliminary investigation fixed-term expressions, the geostatistical binomial logistic models in focus are the models with linear predictors
\begin{align}
\eta_{i} & = \beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{piped\underline{\hspace{0.125cm}}sewer}(x_{i})^{2} (\#eq:0040) \\
& \quad  + \beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}\nonumber
\end{align}
and
\begin{align}
\eta_{i} & = \beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{ln(p\underline{\hspace{0.125cm}}density.k)}(x_{i}) (\#eq:0041) \\
& \quad  + \beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i} \nonumber
\end{align}

The next few sections discuss the models' estimates, and whether the models successfully capture **(a)** unexplained spatial variation, and **(b)** unstructured variation in $Y_{i}$.



\clearpage

## Coefficients

<!--- excludes ln(population density) --->
```{r T0011, fig.pos = 'h!'}
estimations <- readRDS(file = file.path(dirname(getwd()), 'warehouse', 'models', 'nugget', 'blm', 'A', 'estimations.rds'))

coefficients <- estimations$coefficients
coefficients$variable <- c('1', 'piped\\underline{\\hspace{0.125cm}}sewer',
                           'piped\\underline{\\hspace{0.125cm}}sewer$^{2}$', 'elevation.km')

coefficients %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'coef.', 'coef.\nest.',
                                        'lower', 'upper', 's.e.', '$p_{value}$')),
    caption = 'The estimated coefficients of the Binomial geostatistical logistic model with linear predictor $\beta_{0} + \beta_{1}{piped\underline{\hspace{0.125cm}}sewer}(x_{i}) + \beta_{2}{piped\underline{\hspace{0.125cm}}sewer}(x_{i})^{2} + \beta_{3}{elevation.km}(x_{i})  + S(x_{i}) + U_{i}$') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', '', 'confidence\ninterval' = 2, '', '')) %>%
  kableExtra::add_footnote(label = c('coef.: coefficient', 'est.: estimate', 's.e.: standard error'), notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```

<!--- includes ln(population density) --->
```{r T0012, fig.pos = 'h!'}
estimations <- readRDS(file = file.path(dirname(getwd()), 'warehouse', 'models', 'nugget', 'blm', 'B', 'estimations.rds'))

coefficients <- estimations$coefficients
coefficients$variable <- c('1', 'piped\\underline{\\hspace{0.125cm}}sewer',
                           'ln(p\\underline{\\hspace{0.125cm}}density.k)', 'elevation.km')

coefficients %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'coef.', 'coef.\nest.',
                                        'lower', 'upper', 's.e.', '$p_{value}$')),
    caption = 'The estimated coefficients of the \\textit{Case B} Binomial Geostatistical Model') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_header_above(c('', '', '', 'confidence\ninterval' = 2, '', '')) %>%
  kableExtra::add_footnote(label = c('coef.: coefficient', 'est.: estimate', 's.e.: standard error'), notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```





\clearpage

## Residual Spatial Correlation

<!--- excludes ln(population density) --->
```{r F0008, fig.pos = 'h!', out.width = '60%', fig.align = 'center', fig.cap = 'The variogram of the \\textit{Case A} Binomial Geostatistical Model.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'A', 'variogram.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

<!--- includes ln(population density) --->
```{r F0011, fig.pos = 'h!', out.width = '60%', fig.align = 'center', fig.cap = 'The variogram of the \\textit{Case B} Binomial Geostatistical Model.'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'B', 'variogram.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```


\clearpage

## Prevalence Estimates

<!--- excludes ln(population density) --->
```{r F0009, fig.pos = 'h!', out.width = '85%', fig.align = 'center', fig.cap = 'The \\textit{Case A} Binomial Geostatistical Model prevalence estimates and originals'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'A', 'diagonal.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE)
)
```

<!--- includes ln(population density) --->
```{r F0012, fig.pos = 'h!', out.width = '85%', fig.align = 'center', fig.cap = 'The \\textit{Case B} Binomial Geostatistical Model prevalence estimates and originals'}
knitr::include_graphics(
  path = file.path('..', 'warehouse', 'models', 'nugget', 'blm', 'B', 'diagonal.pdf'),
  auto_pdf = getOption(x = "knitr.graphics.auto_pdf", default = TRUE),
  dpi = NULL,
  error = getOption(x = "knitr.graphics.error", default = TRUE))
```


### Bias & RMSE

<!--- excludes ln(population density) --->
```{r T0013}
discrepancies <- estimations$discrepancies
discrepancies <- cbind(variable = row.names(discrepancies), discrepancies)
row.names(discrepancies) <- NULL

discrepancies %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'Bias', 'RMSE')),
    caption = 'Bias and root mean square error w.r.t. the \\textit{Case A} Binomial Geostatistical Model') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```

<!--- includes ln(population density) --->
```{r T0014}
discrepancies <- estimations$discrepancies
discrepancies <- cbind(variable = row.names(discrepancies), discrepancies)
row.names(discrepancies) <- NULL

discrepancies %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    digits = 3,
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('', 'Bias', 'RMSE')),
    caption = 'Bias and root mean square error w.r.t. the \\textit{Case B} Binomial Geostatistical Model') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l") %>%
  kableExtra::column_spec(column = 3, latex_column_spec = "l")
```
