```{r}
sys.source(file = 'R/descriptions/STH.R', envir = knitr::knit_global())
```

# Background

A background reading/research discussion of

* Strategies that have guided STH (soil transmitted helminth) field survey area selection.
* Methods for predicting STH prevalence.
* The selection of fields for STH surveys.
* The biology, ecology, and transmission dynamics of geohelminths.

\vspace{25pt}

Herein, discuss helminthiases, and particularly focus on soil-transmitted helminth infections. School age children are children in the ae range 5 -- 14 years [@Montresor2011]

```{r}
terms <- STH()

terms %>%
  dplyr::mutate_all(linebreak) %>%
  kableExtra::kable(
    format = 'latex',
    booktabs = TRUE,
    escape = FALSE,
    linesep = '\\addlinespace',
    col.names = kableExtra::linebreak(c('species', 'name')),
    caption = 'The soil-transmitted helminth\\\\species in focus') %>%
  kableExtra::kable_styling(position = 'center', latex_options = 'hold_position', font_size = 7, full_width = FALSE) %>%
  kableExtra::add_footnote(label = '', notation = 'none')  %>%
  kableExtra::column_spec(column = 1, italic = TRUE, latex_column_spec = "p{0.18\\\\linewidth}l") %>%
  kableExtra::column_spec(column = 2, latex_column_spec = "l")
```


\vspace{25pt}

There have been a variety of studies about the effects, or otherwise, of environmental factors, e.g., temperature,  rainfall, altitude, soil moisture, relative atmospheric humidity, etc., on geohelminths prevalence spatial patterns [Brooker 2006]

\vspace{25pt}

Under certain conditions eggs may not embryonate
* low humidity, $< 50%$: Ascaris lumbricoides, Trichuris trichiura

